---
title: "EDA"
author: "Maya Zhao"
date: '2022-05-25'
output: html_document
---

Attribute Information:

Date Time: Each ten minutes.
Temperature: Weather Temperature of Tetouan city.
Humidity: Weather Humidity of Tetouan city.
Wind Speed of Tetouan city.
general diffuse flows
diffuse flows
power consumption of zone 1 of Tetouan city.
power consumption of zone 2 of Tetouan city.
power consumption of zone 3 of Tetouan city.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=15, fig.height=8)
```

```{r}
rm(list = ls()) #initialization
```


```{r}
library(tidyverse)
library(data.table)
library(lubridate)
library(dplyr)
library(lubridate)
library(forecast)
library(TSA)
library(xts)
library(zoo)
library(Metrics)
library(ggplot2)
library(gplots)
library(RColorBrewer)
```

```{r}
# read the real data on July 1 for test
df <- read.csv("Tetuan City power consumption.csv")
colnames(df)
```

```{r}
# check for missing values
sum(is.na(df$Zone1))
```


```{r}
colnames(df)[7] <- "Zone1"
colnames(df)[8] <- "Zone2"
colnames(df)[9] <- "Zone3"
date <- seq(from=as.POSIXct("2017-01-01 00:00", format = "%Y-%m-%d %H:%M"), length.out = nrow(df), by = "10 min")
df$DateTime <- date #df(df$DateTime, format="%m/%d/%Y %H:%M")
#attr(df$DateTime, "tzone") <- "Africa/Casablanca"
df_xts <- xts(df[,7], date) 
```



## EDA

### check stationarity 
```{r}
# ts plot and acf plot
var = colnames(df)
var = var[-1]
for(i in var) {       
  ts.plot(df[[i]], ylab=i)
  acf(df[[i]], ylab = paste0(i," ACF"), main="")
}
```



```{r}
library(tseries)
for(i in var) {       
  print(adf.test(df[[i]])) # detects stationarity
  print(kpss.test(df[[i]],null = "Trend")) # detects trend stationarity
}
```


```{r}
#perform frequency decomposition on our data to identify the strongest frequency signal
p <- periodogram(df$Zone1, xlim = c(0,0.06))
max_freq <- p$freq[which.max(p$spec)]
seasonality <- 1/max_freq
seasonality
```

```{r}
#now let's find the remaining top frequencies
max_freqs <- p$freq[p$spec > 50000000000]
max_freqs
```

As noted above, there are certain frequencies that appear. The first two frequencies are so close to zero that they bear no meaning (frequency of 0 means an event does not repeat. We may get additional insights if we could obtain several years' worth of data)

```{r}
seasonalities <- 1/max_freqs
seasonalities
```

As mentioned previously, the first two seasonalities are associated with frequencies that are close to zero and bear no meaning. Seasonalities close to 144 are for daily seasonality (six 10-minute periods per hour times 24 = 144). Seasonalities close to 72 are for semi-daily seasonality (12 hours, six 10-minute periods per hour times 12 = 72). Seasonalities close to 48 are for 8-hour frequencies (six 10-minute periods per hour times 8 = 48.)

```{r}
plot(df$Zone1, type = "l", xlab = "Time", ylab = "Watt Hours", main = "Power Consumption of Tetouan City", col = "#4F94CD")
lines(df$Zone2, col = "#708090")
lines(df$Zone3, col = "#EEE5DE")

legend("topleft", c("Zone 1", "Zone 2","Zone 3"),
       lty = c(1,1),
       col = c("#4F94CD","#708090", "#EEE5DE"))

```


### Correlation
```{r}
cor_df <- df[, c(2,3,4,5,6,7)]
cor_mat <- round(cor(cor_df),2)
heatmap.2(cor_mat,Colv=NA,Rowv=NA,col=brewer.pal(9,"Blues"),cellnote=cor_mat,notecol="white", na.color=par("bg"),trace='none',density.info='none', margins = c(5, 5),cexRow=0.5,cexCol=0.5)
```
```{r}
comp_daily <- decompose(ts(df$Zone1,frequency = 6*24))
plot(comp_daily)
```
```{r}
comp_weekly <- decompose(ts(df$Zone1,frequency = 6*24*7))
plot(comp_weekly)
```

```{r}
plot(df_xts['2017-01-01/2017-01-07'], type = "l", xlab = "Time", ylab = "Watt Hours", main = "One Week of Power Consumption of Zone 1 of Tetouan City ", col = "#4F94CD")
```
```{r}
plot(df_xts['2017-01-01'], type = "l", xlab = "Time", ylab = "Watt Hours", main = "One Day of Power Consumption of Zone 1 of Tetouan City ", col = "#4F94CD")
```

```{r}
plot(df_xts['2017-01-01/2017-01-30'], type = "l", xlab = "Time", ylab = "Watt Hours", main = "One Month of Power Consumption of Zone 1 of Tetouan City ", col = "#4F94CD")
```

