---
title: "linear_model"
author: "Kartik Garg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, warning=FALSE, message=FALSE}
library(TSA, warn.conflicts = FALSE)
library(forecast)
library(Metrics)
library(gdata)
library(readxl)
library(tseries)
library(dplyr)
library(reshape)
library(ggplot2)
```

## Linear Model on Tetuan City power consumption

```{r}
df <- read.csv("Tetuan City power consumption.csv")
head(df)
```

### Feature Engineering and additional EDA

Based on the 00EDA, we look at what features can we add for the linear model

Adding hour and day of the week from the DateTime as an independent variable for our model

```{r}
library(lubridate)
df$DateTime <- as.POSIXct(df$DateTime,format="%m/%d/%Y %H:%M")
df <- df[complete.cases(df), ]
df$hour <- as.integer(format(as.POSIXct(df$DateTime,format="%H:%M:%S"),"%H"))
df$minute <- as.integer(format(as.POSIXct(df$DateTime,format="%H:%M:%S"),"%M"))
df$dow <-as.integer(wday(df$DateTime))
```

```{r}
head(df)
```
```{r}
tail(df)
```

EDA

```{r}
library(corrplot)
corrplot(cor(df[,2:12]), title = "Correlation Plot")
```


Splitting our datainto train and test

```{r}
train<-df[df$DateTime >= "2017-01-01" & df$DateTime < "2017-12-01",]
train<-train[complete.cases(train), ]
tail(train)
```

```{r}
test <- df[df$DateTime >= "2017-12-01",]
test<-test[complete.cases(test), ]
tail(test)
```

## Linear Model 

#### 1. Using Temperature

Here we model using only Temprature as independent variable to get a baseline for our model performance.

y = Zone.1.Power.Consumption
x = Temperature

```{r warning=FALSE}
start_time <- Sys.time()
m1 <- lm(Zone.1.Power.Consumption~Temperature, train)
end_time <- Sys.time()
end_time - start_time


summary(m1)
```
Model forecasting

```{r}
lm_preds1 <- forecast(m1, test)
lm_preds1 <- data.frame(lm_preds1)
pred_df <- data.frame(hours=test$DateTime, test_val=test$Zone.1.Power.Consumption, pred_val=lm_preds1$Point.Forecast)
d<- melt(pred_df, id.vars = "hours")

ggplot(d, aes(x=hours, y=value, color=variable)) + 
  geom_point(size=1) + 
  geom_line()
```


```{r}
smape(test$Zone.1.Power.Consumption,lm_preds1$Point.Forecast) 
rmse(test$Zone.1.Power.Consumption,lm_preds1$Point.Forecast) 
```
This will set a baseline for us to build upon

### Using Temperature & hour

y = Zone.1.Power.Consumption
x = Temperature & hour

```{r warning=FALSE}
start_time <- Sys.time()
m1 <- lm(Zone.1.Power.Consumption~Temperature+hour, train)
end_time <- Sys.time()
end_time - start_time

summary(m1)
```
Model forecasting

```{r}
lm_preds2 <- forecast(m1, test)
lm_preds2 <- data.frame(lm_preds2)
pred_df <- data.frame(hours=test$DateTime, test_val=test$Zone.1.Power.Consumption, pred_val=lm_preds2$Point.Forecast)
d<- melt(pred_df, id.vars = "hours")

ggplot(d, aes(x=hours, y=value, color=variable)) + 
  geom_point(size=1) + 
  geom_line()
```


```{r}
smape(test$Zone.1.Power.Consumption,lm_preds2$Point.Forecast) 
rmse(test$Zone.1.Power.Consumption,lm_preds2$Point.Forecast) 
```
### Using Temperature, hour and dow

y = Zone.1.Power.Consumption
x = Temperature & hour

```{r warning=FALSE}
start_time <- Sys.time()
m1 <- lm(Zone.1.Power.Consumption~Temperature+hour+dow, train)
end_time <- Sys.time()
end_time - start_time

summary(m1)
```
Model forecasting

```{r}
lm_preds3 <- forecast(m1, test)
lm_preds3 <- data.frame(lm_preds3)
pred_df <- data.frame(hours=test$DateTime, test_val=test$Zone.1.Power.Consumption, pred_val=lm_preds3$Point.Forecast)
d<- melt(pred_df, id.vars = "hours")

ggplot(d, aes(x=hours, y=value, color=variable)) + 
  geom_point(size=1) + 
  geom_line()
```


```{r}
smape(test$Zone.1.Power.Consumption,lm_preds3$Point.Forecast) 
rmse(test$Zone.1.Power.Consumption,lm_preds3$Point.Forecast) 
```


## Evaluation

```{r}
Models <- c("Linear regression on temp", "Linear regression on temp and hour",
            "Linear regression on temp, hour and dow")

sMAPE <- c(smape(test$Zone.1.Power.Consumption,lm_preds1$Point.Forecast),
smape(test$Zone.1.Power.Consumption,lm_preds2$Point.Forecast),
smape(test$Zone.1.Power.Consumption,lm_preds3$Point.Forecast))

RMSE <- c(rmse(test$Zone.1.Power.Consumption,lm_preds1$Point.Forecast),
rmse(test$Zone.1.Power.Consumption,lm_preds2$Point.Forecast),
rmse(test$Zone.1.Power.Consumption,lm_preds3$Point.Forecast))

res.df<- data.frame(Models, sMAPE, RMSE)
res.df
```

Based on the results we can see that the best model is the one with temp and hour

