---
title: "tbats_model"
author: "Kartik Garg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, warning=FALSE, message=FALSE}
library(TSA, warn.conflicts = FALSE)
library(forecast)
library(Metrics)
library(gdata)
library(readxl)
library(tseries)
library(dplyr)
library(reshape)
library(ggplot2)
```

## TBATS Model on Tetuan City power consumption

```{r}
df <- read.csv("Tetuan City power consumption.csv")
head(df)
```

Splitting into train and test

```{r}
df$DateTime <- as.POSIXct(df$DateTime,format="%m/%d/%Y %H:%M")
df2<-df[df$DateTime >= "2017-01-01" & df$DateTime < "2017-12-01",]
df2<-df2[complete.cases(df2), ]
history <- data.frame(ds = df2$DateTime, y = df2$Zone.1.Power.Consumption)
tail(history)
```

```{r}
test <- df[df$DateTime >= "2017-12-01",]
test<-test[complete.cases(test), ]
```


## TBATS with varying seasonality


We will build the model with varying seasonalities and cross check the results

### Daily

```{r}
start_time <- Sys.time()
m1 <- auto.arima(ts(history$y, frequency = 144))
end_time <- Sys.time()
end_time - start_time
```


```{r}
start_time <- Sys.time()
m1 <- tbats(ts(history$y, frequency = 144), seasonal.periods = 144)
end_time <- Sys.time()
end_time - start_time

comp <-tbats.components(m1)
plot(comp)
```

seasonal components for the daily seasonality model

```{r}
tbats_preds1 <- forecast(m1, h=nrow(test))
plot(tbats_preds1)
```

Let's plot same with our test data and check the results now


```{r}
tbats_preds1 <- data.frame(tbats_preds1)
pred_df <- data.frame(hours=test$DateTime, test_val=test$Zone.1.Power.Consumption, pred_val=tbats_preds1$Point.Forecast)
d<- melt(pred_df, id.vars = "hours")

ggplot(d, aes(x=hours, y=value, color=variable)) + 
  geom_point(size=1) + 
  geom_line()
```

Looks very accurate!


```{r}
smape(test$Zone.1.Power.Consumption,tbats_preds1$Point.Forecast) 
rmse(test$Zone.1.Power.Consumption,tbats_preds1$Point.Forecast) 
```
Great results for sMAPE and RMSE

### Weekly

```{r}
start_time <- Sys.time()
m1 <- tbats(ts(history$y, frequency = 144*7), seasonal.periods = 144*7)
end_time <- Sys.time()
end_time - start_time

comp <-tbats.components(m1)
plot(comp)
```

These are the seasonal components for the weekly seasonality model


```{r}
tbats_preds2 <- forecast(m1, h=nrow(test))
plot(tbats_preds2)
```

Not too bad, let's check it with our test data

```{r}
tbats_preds2 <- data.frame(tbats_preds2)
pred_df <- data.frame(hours=test$DateTime, test_val=test$Zone.1.Power.Consumption, pred_val=tbats_preds2$Point.Forecast)
d<- melt(pred_df, id.vars = "hours")
pred_df
ggplot(d, aes(x=hours, y=value, color=variable)) + 
  geom_point(size=1) + 
  geom_line()
```

```{r}

smape(test$Zone.1.Power.Consumption,tbats_preds2$Point.Forecast)
rmse(test$Zone.1.Power.Consumption,tbats_preds2$Point.Forecast)
```

## Evaluation

Let's analyze the results

```{r}
Models <- c("Model daily", "Model weekly")

sMAPE <- c(smape(test$Zone.1.Power.Consumption,tbats_preds1$Point.Forecast),
smape(test$Zone.1.Power.Consumption,tbats_preds2$Point.Forecast))

RMSE <- c(rmse(test$Zone.1.Power.Consumption,tbats_preds2$Point.Forecast),
rmse(test$Zone.1.Power.Consumption,tbats_preds2$Point.Forecast))

res.df<- data.frame(Models, sMAPE, RMSE)
res.df
```
From the results of the model we can see that the daily model outperforms the weekly model